---
title: "Using an Extreme Gradient Boosting Learner for Mapping Hydrogeochemical Parameters in Germany"
subtitle: "Current State of the Project"
author: "Maximilian Nölscher, Stefan Broda"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  bookdown::html_document2:
    theme: lumen
    highlight: tango
    toc: TRUE
    toc_float: TRUE
    collapse: FALSE
    toc_depth: 3
    collapsed: FALSE
    smooth_scroll: FALSE
    number_sections: TRUE
    self_contained: yes
    df_print: paged
    #code_folding: show
bibliography: vEGU_display_material.bib
nocite: '@*'
biblio-style: "apalike"
link-citations: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(fig.align="center")
options(tidyverse.quiet = TRUE)
```

```{r}
library(skimr)
library(visdat)
library(targets)
library(mapview)
library(knitr)
library(sf)
library(ggExtra)
library(patchwork)
library(tidyverse)
```

```{r, echo = FALSE}
source("R/processing_functions.R")
source("R/markdown.R")
source("R/plot_functions.R")
source("R/constants.R")
```

<br /> This study is part of the [MACRO
project](https://www.bgr.bund.de/EN/Themen/Wasser/Projekte/laufend/F+E/Macro/macro_projektbeschr_en.html?nn=1548270).
This project will end approximately in the summer of 2023. Briefly, its
goals are to explore the application of machine learning for
mapping/regionalization in the field of hydrogeology in general. <br />
<br />

<p align="center">

<a href="https://www.bgr.bund.de/EN/Themen/Wasser/Projekte/laufend/F+E/Macro/macro_projektbeschr_en.html?nn=1548270">
<img src="logo/MACRO_logo.svg"/> </a>

</p>

<br /> <br /> <br /> <br /> <br />

# Abstract

Information on the spatial distribution of hydrogeochemical parameters
is crucial for decision making. Machine learning based methods for the
mapping of hydrogeochemical parameter concentrations have been already
studied for many years to evolve from deterministic and geostatistical
interpolation methods. However, the reflection of all relevant processes
that the target variables depend on is often difficult to achieve,
because of the mostly insufficient determination and/or availability of
features. This is especially true if you limit yourself to freely
accessible data.

In this study, we apply an extreme gradient boosting learner (XGBoost)
to map major ion concentrations across Germany. The training data
consists of water samples from approximately 35K observation wells
across Germany and a wide range of environmental data as predictors. The
water samples were collected between the 1950s and 2005 at
anthropogenically undisturbed locations.

The environmental data includes hydrogeological units and parameters,
soil type, lithology, digital elevation model (DEM) and DEM derived
parameters etc. The values of these features at the respective water
sample location were extracted on the basis of a polygon, approximately
representing the area that has an impact on the target variable (ion
concentration). For a comparison, different polygon shapes are used.

# Workflow

The workflow from data preprocessing to model evaluation is
schematically shown in fig. \@ref(fig:flowchart). The single steps are
described more in detail in the following sections. The work of this
study is still in progress. The preprocessing of the hydrogeochemical
background values dataset is described in [Hydrogeochemical Parameters],
the feature extraction in section [Feature Extraction], the modelling
pipeline in [Model Training] and the evaluation can be found in
[Evaluation].

All data processing and modelling was done using the R programming
language using multiple packages (see chapter [References]).

```{r flowchart, fig.cap="flowchart", out.width="50%"}
knitr::include_graphics("inkscape_figures/flowchart.svg")
```

# Data

## Hydrogeochemical Parameters

```{r}
original_data <- tar_read(back_vals_2005_clean_sf)
```

The target variables of the training data being used in this study is
based on a dataset containing initially approximately
`r as.character(round(nrow(original_data), -3))` measurements of
hydrogeochemical parameters from groundwater samples predominantly taken
during the second half of the 20th century until
`r original_data %>% pull(date) %>% max(na.rm = TRUE) %>% lubridate::year()`.

Before preprocessing the data, we can have a quick view on some
attributes of the dataset: sample date and sample depth.

**Distribution of the sample date** <br />

```{r sampledatedistribution, out.width="70%", fig.cap="Distribution of sample date"}
tar_read(sampledatedistribution_current_plot)
```

Whereas the earliest samples are from the year
`r tar_read(back_vals_2005_clean_sf) %>% pull(date) %>% min() %>% lubridate::year()`,
most samples were taken in the second half of the 20^th^ century.

<br /> **Distribution of the sample depth** <br />

The sample depth was calculated as mean depth of the screen top and
bottom if it was provided. Thus, it reflects the screen center depth
below ground level of an observation well. For the model training, the
sample depth was used as feature (predictor variable).

```{r sampledepthdistribution, out.width="70%", fig.cap="Distribution of sample depth"}
tar_read(sampledepthdistribution_current_plot)
```

<br />

The number of samples used for training the models is reduced due to the
following steps:

### Preprocessing

<br /> **Intersection with Study Area** <br />

The sample locations were intersected with the administrative border of
Germany.

<br /> **Aggregation of multiple measurements per sample site** <br />

Some of the sample sites have multiple measurements over time which were
aggregated by calculating the mean. The distribution of the number of
measurements per sample site after applying the previous processing
steps is shown in figure \@ref(fig:measurementdistribution).

```{r measurementdistribution, out.width="70%", fig.cap="Distribution of multiple measurements per sample site"}
tar_read(measurementdistribution_current_plot)
```

```{r}
target_data <-
  tar_read(back_vals_filter_allpositive_sf) %>%
  as_tibble() %>%
  select(station_id, sample_depth, all_of(tar_read(parameters_to_model))) %>%
  filter_all(any_vars(complete.cases(.)))
```

<br /> **Selection of parameters as target variables** <br />

From all measured parameters, the ten ions with the most samples were
selected as target variables to be modelled
(`r tar_read(parameters_to_model) %>% parameter_pretty_markdown() %>% word(1) %>% str_c(collapse = ", ")`;
see fig. \@ref(fig:samplesparameters)). Across all these parameters and
after all preprocessing steps, `r nrow(target_data)` samples and
`r ncol(target_data)` columns, 1 for the station ID (`station_id`), 1
for the sample depth (`sample_depth`) and `r ncol(target_data) - 2` for
each target variable remain for the model training.

```{r samplesparameters, fig.height=8, fig.cap="Samples per hydrogeochemical."}
tar_read(samplesparameters_current_plot)
```

### Location

The locations of the sample sites used for modelling are shown in
\@ref(fig:locations) as the number of sample sites per hexagon. The
spatial distribution of sampling locations is unbalanced with regions
that have few locations and regions with a high density. The latter are
mainly concentrated around larger cities in Germany such as Berlin,
Hamburg, Frankfurt. The eastern and northern areas of Germany also
generally have more sampling sites compared to southern or central
Germany.

```{r locations, fig.cap="Sample site locations"}
tar_read(locations_current_plot)
```

### Data Summary

The resulting dataset steps is shown in table \@ref(tab:targetdatahead)
as an example.

```{r targetdatahead}
target_data %>% 
  drop_na(everything()) %>% 
  select(-sample_depth) %>% 
  slice(1:3) %>% 
  knitr::kable(caption = "First three rows of the dataset containing the target variables")
```

<br />
<br />

The distribution of target values as violin chart is shown in figure
\@ref(fig:violindistribution). From the distribustions of some of the parameters, it can be inferred that values referring to the lower detection limit are still present in the dataset (e.g. see high concentration of potassium concentrations below 1E-4 mg/L). This values must be removed. 

```{r violindistribution, out.width="70%", fig.cap="Distribution of the target variable values"}
tar_read(plot_xgb_violin_distribution_targets)
```



## Features

```{r}
data_features <- 
  tar_read(data_features_depth_added)
```

### Feature Extraction

In addition to this dataset, geophysical attributes were extracted from
other spatial data sources (see the following list) and used as
features:

-   [Geological
    Map](https://www.bgr.bund.de/DE/Themen/Sammlungen-Grundlagen/GG_geol_Info/Karten/Deutschland/GUEK200/guek200_inhalt.html)
    [@bgr__sgd_guk250_2021]
-   [Hydrogeological
    Map](https://www.bgr.bund.de/DE/Themen/Wasser/Projekte/laufend/Beratung/Huek200/huek200_projektbeschr.html)
    [@bgr__sgd_huk250_2019]
-   [Seepage
    Rate](https://www.bgr.bund.de/DE/Themen/Boden/Produkte/Karten/Downloads/Berechnung_Sickerwasserrate.pdf?__blob=publicationFile&v=4)
    [@bgr_swr1000_2003]
-   [Groundwater
    Recharge](https://www.bgr.bund.de/DE/Themen/Wasser/Projekte/abgeschlossen/Beratung/Had/Was_had_abb_gwn1000.html?nn=1546102)
    [@bgr_gwn1000_2019]
-   [Hydrogeological
    Units](https://www.bgr.bund.de/DE/Themen/Wasser/Projekte/abgeschlossen/Beratung/Hyraum/hyraum_projektbeschr.html)
-   [Longterm Mean
    Temperature](https://opendata.dwd.de/climate_environment/CDC/grids_germany/multi_annual/air_temperature_mean/BESCHREIBUNG_gridsgermany_multi_annual_air_temperature_mean_7100_de.pdf)
    [@noauthor_vieljahrige_nodate]
-   [Longterm Mean
    Precipitation](https://opendata.dwd.de/climate_environment/CDC/grids_germany/multi_annual/precipitation/BESCHREIBUNG_gridsgermany_multi_annual_precipitation_6190_de.pdf)
-   Multiorder Hydrologic Position (generated for this study after
    @belitz_multiorder_2019, not yet published)
    [@noauthor_raster_nodate]
-   [Soil
    units](https://www.bgr.bund.de/DE/Themen/Boden/Bilder/Bod_BGL5000_g.html)
    [@bgr_bgl5000_2014]
-   [Land Use and Land
    Cover](https://land.copernicus.eu/pan-european/corine-land-cover)
    [@_european_union_corine_2018]
-   [Elevation](https://land.copernicus.eu/imagery-in-situ/eu-dem/eu-dem-v1.1)
    [@_european_union_eu-dem_nodate]
-   [Slope](https://land.copernicus.eu/imagery-in-situ/eu-dem/eu-dem-v1-0-and-derived-products/slope)
    [@_european_union_slope_nodate]
-   [Aspect](https://land.copernicus.eu/imagery-in-situ/eu-dem/eu-dem-v1-0-and-derived-products/aspect)
    [@_european_union_aspect_nodate]

The features were extracted based on a 1km buffer around the sample site
as approximated groundwater contributing area as follows (see figure
\@ref(fig:interactiveschematiclulcfeatureextractionmap)):

-   Categorical features: the share of each class in the buffer area was
    calculated. At the same time, this leads to an encoding as numerical
    features.

-   Numerical data: the mean for the buffer was calculated.

    A basic evaluation of the impact of buffer sizes was conducted in
    Knoll et al. (2019).

    The feature extraction leads to a total number of
    `r data_features %>% ncol() %>% magrittr::subtract(1)` features.

```{r interactiveschematiclulcfeatureextractionmap, fig.cap="Example of feature extraction based on circular buffer around sample sites (red) (e.g. the land use and land cover data as shown here)"}
knitr::include_graphics("plots_from_mohp_data_preparation/interactive_schematic_lulc_feature_extraction_map.svg")
```

The first three rows of the dataset containing the features is shown in
table \@ref(tab:featuresdatahead) as an example.


```{r featuresdatahead}
data_features %>% 
  drop_na(everything()) %>% 
  slice(1:3) %>% 
  knitr::kable(caption = "First three rows of the dataset containing the features") %>% 
  kableExtra::scroll_box(width = "800px")
```

# Model Training

All the data preprocessing all modelling was done using the R
programming language and the `tidymodels` package. The modelling
pipeline includes the following steps:

-   Normalization (standardization) of all features

-   Removal of highly correlated and/or sparse features

-   80/20 train-test split

-   5-fold cross-validation and stratification of the target variable
    (no spatial cross-validation yet implemented)

-   Hyperparameter tuning of the model parameters

    -   `min_n`: minimum sum of instance weight (hessian) needed in a
        child [@chen_xgboost_2020]
    -   `tree_depth`: maximum depth of a tree [@chen_xgboost_2020]
    -   `learn_rate`: learning rate [@chen_xgboost_2020]
    -   `loss_reduction`: minimum loss reduction required to make a
        further partition on a leaf [@chen_xgboost_2020] node of the
        tree

    for a XGBoost learner (number of trees fixed to 1000) with 50
    parameter combinations using a parameter space filling grid.

# Results

## Hyperparameter Tuning {.tabset}

The following table summarizes the hyperparameters of the best model
configuration after the tuning process.

```{r besthyperparamssummary, echo = FALSE, results = "asis"}
model_params <- tar_read(xgboost_model_final_params)

model_params %>% 
  set_names(tar_read(parameters_to_model)) %>% 
  imap(~mutate(.x, parameter = .y, .before = 1)) %>% 
  reduce(bind_rows) %>% 
  mutate(parameter = parameter_pretty_rmarkdown(parameter, with_unit = FALSE)) %>% 
  kable() %>% 
  kableExtra::kable_styling(full_width = TRUE)
```

## Evaluation

*Disclaimer: As these results are preliminary and this study is still
work in progress, the plain performance metrics are provided without any
further statements or interpretations.*

### Performance {.tabset}

```{r, echo = FALSE, results = "asis"}
plot_observed_vs_predicted <- tar_read(plot_xgb_observed_vs_predicted)
plot_residuals_vs_predicted <- tar_read(plot_xgb_residuals_vs_predicted)

names_pretty <-
  plot_observed_vs_predicted %>%
  names() %>%
  parameter_pretty_markdown(with_unit = FALSE)

for (i in seq_along(names_pretty)) {
  create_section_results(plot_observed_vs_predicted[[i]],
                         title = names_pretty[[i]],
                         level = 4)
}
```


### Feature Importance {.tabset}

```{r, echo = FALSE, results = "asis"}
plot_feature_importance <- tar_read(plot_xgb_feature_importance)

plot_feature_importance %>%
  names() %>%
  parameter_pretty_markdown(with_unit = FALSE) %>% 
  map2(plot_feature_importance, ~create_section(.y, .x, 4))
```

```{r, echo=FALSE}
plot_feature_importance %>% 
  map(function(x) {
    x %>% 
      ggplot_build() %>% 
      chuck("plot") %>% 
      chuck("data") %>% 
      mutate(rank = row_number())
  }) %>% 
  reduce(left_join, by = "Variable") %>% 
  drop_na(everything()) %>% 
  select(-contains("Importance")) %>% 
  pivot_longer(-Variable) %>% 
  select(-name) %>% 
  group_by(Variable) %>% 
  summarise(total_rank = sum(value)) %>% 
  arrange(total_rank) %>% 
  slice(1:5) %>% 
  pull(Variable)
```


# Conclusion and Outlook

This study aims on setting up a benchmark model as a basis for further
development of more sophisticated machine learning models. The
relatively simple model approach shows that the model performance varies
strongly between the target variables indicating that further features
are required to reflect the geophysical characteristics which play a
role in the processes that drive the concentration of the target
variable. Further steps to increase this preliminary model setup are:

-   Inclusion of more and relevant geophysical attributes as features
-   Evaluation of different feature extraction/engineering methods
-   Set up of a model ensemble to approximate a XGBoost multi-output
    model and model all parameters in a single model
-   Implement spatial cross-validation for more realistic model
    performance estimation
-   Evaluate deep learning approaches

# References
